Popularity, Mixed Matchings, and Self-duality

### 2017-03-27 CO750-ApproximationAlgorithmsTopics/assignments/presentation.tex
Writing up slides

\begin{document}

\maketitle

@@

\begin{frame}
\frametitle{Extended Formulation}

\end{frame}
\section{Integrality of $P_G$ in a special case}

\section{Half-Integrality of $P_G$ in any bipartite instance}
%maybe omit this as next section generalizes and uses same ideas
\section{Half-Integrality of $P_G$ in a roommates instance}

>>>>

\begin{document}
\setbeamercolor{background canvas}{bg=white}
\maketitle

@@

\begin{frame}
\frametitle{Extended Formulation}
Fix a popular fractional matching $\alert{y}$. Then $\alert{\Delta(x,y)}$ is a linear function in $\alert{x}$. Since $\alert{y}$ is popular $\alert{\Delta(x,y) \leq 0}$ with equality when $\alert{x}$ is popular.
\begin{columns}[T]
\begin{column}{.48\textwidth}
\begin{align*}
\max &\Delta(x,y) \\
\text{s.t.} \sum_{e \in \delta(u)} x_e &= 1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E
\end{align*}
\end{column}
\begin{column}{.48\textwidth}
\begin{align*}
\min &\sum_{u \in V(G)} \alpha_u  \\
\text{s.t.} \alpha_a + \alpha_b &\geq vote_a(b,y) \\&+ vote_b(a,y) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u), y) &\forall u \in V(G).
\end{align*}
\end{column}
\end{columns}
Where $\alert{\ell(u)}$ is a last choice vertex for each $\alert{u}$ so that each vertex is fully matched in any fractional matching.

We call $\alert{\alpha}$ such that $\alert{\sum_{u} \alpha_u = 0}$ a witness to the popularity of $\alert{y}$.
\end{frame}

\begin{frame}
\frametitle{Extended Formulation}
From the primal-dual pair we obtain the extended formulation $\alert{(LP)}$:
\begin{align*}
\min &\sum_{u \in V(G)} \alpha_u \\
\text{s.t.} \alpha_a + \alpha_b &\geq vote_a(b,x) + vote_b(a,x) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u),x) &\forall u \in V(G) \\
\sum_{e \in \delta(u)} x_e &=1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E.
\end{align*}
Optimal value is $0$. The optimal solutions are $\text{\alert{popular matchings}}$ and their $\text{\alert{witnesses}}$ (see Popular Mixed Matchings by Kavitha et al. for details).
\end{frame}

\begin{frame}
\frametitle{Extended Formulation}
Let $\alert{P'_G}$ be the optimal solutions to $\alert{(LP)}$. Then $\alert{P'_G}$ is described by the set of $\alert{(x,\alpha)}$ such that:
\begin{align*}
\sum_{u \in V(G)} \alpha_u &= 0 \\
\text{s.t.} \alpha_a + \alpha_b &\geq vote_a(b,x) + vote_b(a,x) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u),x) &\forall u \in V(G) \\
\sum_{e \in \delta(u)} x_e &=1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Self-Duality}
It is straightforward to see that $\alert{(LP)}$ is self-dual (exercise). This implies, by $\text{\alert{complementary slackness}}$, that the covering constraint for $\alert{ab}$ is tight when $\alert{x_{ab} > 0}$ for any popular fractional matching $\alert{x}$.

$\textbf{Lemma 1}$: For any $\alert{(x,\alpha) \in P'_G}$, if $\alert{x_{ab} > 0}$ then:
$$\alpha_a + \alpha_b = vote_a(b,x) + vote_b(a,x).$$ 
\end{frame}
\section{Integrality of $P_G$ in a special case}
\begin{frame}
\frametitle{Objective}
$\textbf{Theorem 1:}$ If $\alert{G =(A\cup B, E)}$ is $\text{\alert{bipartite}}$ and admits a $\text{\alert{perfect stable matching}}$ then $\alert{P_G}$ is integral.

From prior talks: $\text{\alert{stable matchings}}$ are $\text{\alert{minimum size popular matchings}}$.

So the hypothesis says $\text{\alert{popular matching}}$ is perfect and, in the fractional setting, for each $\alert{u\in A \cup B}$, $\alert{x_{u\ell(u)} = 0}$.

To prove $\textbf{Theorem 1}$ we will write $\alert{(x,\alpha^x) \in P_G'}$ as a convex combination of $\text{\alert{popular matchings}}$.
\end{frame}

\begin{frame}
\frametitle{Bounding $\alpha^x$}
$\textbf{Lemma 2:}$ $\alert{\alpha^x \in [-1,1]}$.

$\textbf{Proof:}$ First
$$\alpha^x_u \geq vote_u(\ell(u), x) = -1 \text{(since }x_{u,\ell(u)} = 0\text{).}$$
Now let $\alert{v}$ be the least preferred neighbour of $\alert{u}$ such that $\alert{x_{uv} >0}$. Then
$$vote_u(v,x) = -vote_u(x,v) = -(1-x_{uv}).$$
Also
$$vote_v(u,x) = \sum_{u':u' <_v u} x_{u'v} - \sum_{u':u' >_v u} x_{u'v} = 1-x_{uv}.$$
So by $\textbf{Lemma 1}$
$$\alpha_u + \alpha_v = vote_u(v,x) + vote_v(u,x) \leq -(1-x_{uv}) + (1-x_{uv}) = 0.$$
Since $\alert{\alpha_v \geq -1}$ this implies $\alert{\alpha_u \leq 1}$.$\blacksquare$
\end{frame}

\begin{frame}
\frametitle{Constructing the Table: $a \in A$}
\centering
$r_a\cdot 1 + (1-r_a)(-1) = \alpha^x_a$ 
\includegraphics[scale=0.25]{Xa}

For each $\alert{ab \in \delta(a)}$ create a cell of length $\alert{x_{ab}}$ (omit $0$-length cells) and arrange cells in increasing order (w.r.t. $\alert{a}$) to form $\alert{X_a}$.
\end{frame}

\begin{frame}
\frametitle{Constructing the Table: $b \in B$}
\centering
$r_b(-1) + (1-r_b)\cdot 1 = \alpha^x_b$
\includegraphics[scale=0.25]{Xb}

For each $\alert{ab \in \delta(b)}$ create a cell of length $\alert{x_{ab}}$ (omit $0$-length cells) and arrange cells in decreasing order (w.r.t. $\alert{b}$) to form $\alert{X_b}$.
\end{frame}

\begin{frame}
\frametitle{Constructing the Table: Finding Matchings}
We form the table $\alert{T}$ whose rows are the arrays $\alert{X'_u}$ for each $\alert{u \in V(G)}$.

For any $\alert{t \in [0,1)}$ form $\alert{M_t\subseteq E}$ as follows:
\begin{enumerate}
\item Draw a vertical line at distance $\alert{t}$ from left end of $\alert{T}$.
\item For each $\alert{u \in V(G)}$ define $\alert{c_u(t)}$ to be the cell in $\alert{X'_u}$ which the line intersects (or touches the left boundary of).
\item Set $\alert{M_t = \{ uv \in E(G) : u \in V(G) \text{and} v \in c_u(t)\}}$.
\end{enumerate}

We claim that each $\alert{M_t}$ is a $\text{\alert{popular matching}}$.
\end{frame}

\begin{frame}
\frametitle{Brief detour to pick up a Lemma}
$\textbf{Lemma 3}$ For any $\alert{ab \in E(G)}$,
$$ x_{ab} + \sum_{b' <_a b} x_{ab'} + \sum_{a' <_b a} x_{a'b} \leq r_a + (1-r_b).$$
Equality holds if $\alert{x_{ab} > 0}$.

$\textbf{Proof:}$
From the covering constraint for $\alert{P'_G}$:
$$\alpha^x_a + \alpha^x_b \geq \sum_{b' <_a b} x_{ab'} - \sum_{b' >_a b} x_{ab'} + \sum_{a' <_b a} x_{a'b} - \sum_{a' >_b a} x_{a'b}.$$
Substitute $\alert{\sum_{b' >_a b} x_{ab'} = 1- x_{ab} - \sum_{b'<_a b} x_{ab'}}$, and similar for $\alert{\sum_{a' >_b a}x_{a'b}}$. Substitute $\alert{\alpha^x_a = 2r_a -1}$ and $\alert{\alpha^x_b = 1-2r_b}$.

The result follows. $\textbf{Lemma 1}$ yields equality when $\alert{x_{ab}} > 0$.$\blacksquare$
\end{frame}

\begin{frame}
\frametitle{$M_t$ is a matching}
%May take multiple frames
\end{frame}

\begin{frame}
\frametitle{$M_t$ is popular}
%May take multiple frames
\end{frame}
\section{Half-Integrality of $P_G$ in any bipartite instance}
%maybe omit this as next section generalizes and uses same ideas
\section{Half-Integrality of $P_G$ in a roommates instance}
